\LARGE \bf Planning with Attitude

### 2020-12-29 paper/main.tex
tiny edits

        Let $\phi \in \R^3$ be a differential rotation \added{from the group identity to the 
        input of a vector-valued function of a quaternion} 
        $y = h(q): \Q \to \R^p$, such that
        \begin{equation} \label{eq:vector_function}
            y + \delta y = h(L(q) \varphi(\phi)) \approx h(q) +  \nabla h(q) \phi.
        \end{equation}
        We can calculate the Jacobian $\nabla h(q) \in \R^{p \times 3}$ by
        differentiating \eqref{eq:vector_function} with respect to $\phi$, evaluated at
        $\phi = 0$:
        \begin{equation} \label{eq:quat_gradient}
            \nabla h(q) = \pdv{h}{q} L(q) H := \pdv{h}{q} G(q) 
                        = \pdv{h}{q} \begin{bmatrix} 
                            -q_v^T \\ 
                            q_s I_3 + \skewmat{q_v}
                        \end{bmatrix}
        \end{equation}
        where $G(q) \in \R^{4 \times 3}$ is the \textit{attitude Jacobian}, which
        essentially becomes a ``conversion factor'' allowing us to apply results from
        standard vector calculus to the space of unit quaternions. This form is
        particularly useful in practice since $\pdv*{h}{q} \in \R^{p \times 4}$ can be
        obtained using finite differences or automatic differentiation.
        As an aside, although we have used Rodrigues parameters, $G(q)$ is actually the
        same (up to a constant scalar factor) for any choice of three-parameter attitude
        representation.

@@

    \subsection{Results} Figure \ref{fig:wahba_convergence} compares the
    multiplicative Gauss-Newton method with a na\"ive Newton's method in
    which the quaternion is simply projected back onto the unit sphere at
    every iteration. The na\"ive method makes progress initially, but quickly
    stalls. By correctly handling the group structure of unit quaternions,
    the multiplicative method is able to maintain the fast convergence rates
    typical of Newton's method. \added{By comparing our method with the
    global solution obtained from an SVD decomposition
    \cite{markley1999estimate}, we see that our method recovers the globally
    optimal solution within a small number of iterations.}

    \begin{figure}
        \centering
        \includegraphics[width=\columnwidth, height=5cm]{figures/wahba_convergence.tikz}
        \caption{Convergence comparison for Wahba's problem. \added{The error
        is the angle between the current solution and the globally optimal
        solution computed using SVD decomposition.
        The thick line is the average result of 100 trials with randomized
        orientations and measurements. The thin lines are the maximum and
        minimum over all 100 trials.} By modfying Newton's method with the
        methods of section \ref{sec:Quaternion_Calculus}, quadratic
        convergence rates are achieved, while a na\"ive approach stalls after
        only a few iterations.}
        \label{fig:wahba_convergence}
    \end{figure}

>>>>

        Let $\phi \in \R^3$ be a differential rotation \added{applied to the 
        input of a vector-valued function of a quaternion} 
        $y = h(q): \Q \to \R^p$, such that
        \begin{equation} \label{eq:vector_function}
            y + \delta y = h(L(q) \varphi(\phi)) \approx h(q) +  \nabla h(q) \phi.
        \end{equation}
        We can calculate the Jacobian $\nabla h(q) \in \R^{p \times 3}$ by
        differentiating \eqref{eq:vector_function} with respect to $\phi$, evaluated at
        $\phi = 0$:
        \begin{equation} \label{eq:quat_gradient}
            \nabla h(q) = \pdv{h}{q} L(q) H := \pdv{h}{q} G(q) 
                        = \pdv{h}{q} \begin{bmatrix} 
                            -q_v^T \\ 
                            q_s I_3 + \skewmat{q_v}
                        \end{bmatrix}
        \end{equation}
        where $G(q) \in \R^{4 \times 3}$ is the \textit{attitude Jacobian}, which
        essentially becomes a ``conversion factor'' allowing us to apply results from
        standard vector calculus to the space of unit quaternions. This form is
        particularly useful in practice since $\pdv*{h}{q} \in \R^{p \times 4}$ can be
        obtained using finite differences or automatic differentiation.
        As an aside, although we have used Rodrigues parameters, $G(q)$ is actually the
        same (up to a constant scalar factor) for any choice of three-parameter attitude
        representation.

@@

    \subsection{Results} Figure \ref{fig:wahba_convergence} compares the
    multiplicative Gauss-Newton method with a na\"ive Newton's method in
    which the quaternion is simply projected back onto the unit sphere at
    every iteration. The na\"ive method makes progress initially, but quickly
    stalls. By correctly handling the group structure of unit quaternions,
    the multiplicative method is able to maintain the fast convergence rates
    typical of Newton's method. \added{By comparing our method with the
    global solution obtained from a singular-value decomposition
    \cite{markley1999estimate}, we see that our method recovers the globally
    optimal solution within a small number of iterations.}

    \begin{figure}
        \centering
        \includegraphics[width=\columnwidth, height=5cm]{figures/wahba_convergence.tikz}
        \caption{Convergence comparison for Wahba's problem. \added{The error
        is the angle between the current solution and the globally optimal
        solution computed using a singular-value decomposition.
        The thick line is the average result of 100 trials with randomized
        orientations and measurements. The thin lines are the maximum and
        minimum over all 100 trials.} By modfying Newton's method with the
        methods of section \ref{sec:Quaternion_Calculus}, quadratic
        convergence rates are achieved, while a na\"ive approach stalls after
        only a few iterations.}
        \label{fig:wahba_convergence}
    \end{figure}
